---
title: "Homework_6"
author: "Gretchen Saveson"
date: "2/19/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1
```{r}
library(ggplot2) # for graphics
library(MASS) # for maximum likelihood estimation


# quick and dirty, a truncated normal distribution to work on the solution set

z <- rnorm(n=3000,mean=0.2)
z <- data.frame(1:3000,z)
names(z) <- list("ID","myVar")
z <- z[z$myVar>0,]
str(z)
summary(z$myVar)

# plot histogram of data
# y axis is rescaled so that the area under the curve equals one. 
p1 <- ggplot(data=z, aes(x=myVar, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) 
print(p1)

# add empirical density curve
# this adds a curve that id fitted to the data to smooth the shape of the histogram. It does not assume any distribution. 
p1 <-  p1 +  geom_density(linetype="dotted",size=0.75)
print(p1)

# get maximum likelihood parameters for normal

normPars <- fitdistr(z$myVar,"normal")
print(normPars) # rpints mean and standard deviation
str(normPars)
normPars$estimate["mean"] # note structure of getting a named attribute

# plot normal probability density
# first, assign mean and sd to variables
meanML <- normPars$estimate["mean"]
sdML <- normPars$estimate["sd"]

# x values are specified from 0 to the length of variable of interest.
xval <- seq(0,max(z$myVar),len=length(z$myVar))

# dnorm() is called inside ggplot's "stat_function" to generate the probability density for the normal distribution. 
 stat <- stat_function(aes(x = xval, y = ..y..), fun = dnorm, colour="red", n = length(z$myVar), args = list(mean = meanML, sd = sdML))
 p1 + stat
# The red curve shows the best-fitting normal dist. It has a biased mean because these data have no negative values. The mean of the fitted model is much higher (~0.9) than the mean of the original data (0.2). The normal dist doesn't work well in this instance. 
 
 
# plot exponential probability density
# fitdistr() is used to fit exponential curve to data 
expoPars <- fitdistr(z$myVar,"exponential")
rateML <- expoPars$estimate["rate"]

stat2 <- stat_function(aes(x = xval, y = ..y..), fun = dexp, colour="blue", n = length(z$myVar), args = list(rate=rateML))
 p1 + stat + stat2

# plot uniform probability density
# fitdistr() is not needed. Only two parameters are min and max of data. 
 stat3 <- stat_function(aes(x = xval, y = ..y..), fun = dunif, colour="darkgreen", n = length(z$myVar), args = list(min=min(z$myVar), max=max(z$myVar)))
 p1 + stat + stat2 + stat3

# plot gamma probability density
# fitdistr() is used again here. 
 gammaPars <- fitdistr(z$myVar,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]

stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="brown", n = length(z$myVar), args = list(shape=shapeML, rate=rateML))
 p1 + stat + stat2 + stat3 + stat4
 
# plot beta probability density
#it is plotted in its own plot because data must be rescaled to be between 1 and 0.
 pSpecial <- ggplot(data=z, aes(x=myVar/(max(myVar + 0.1)), y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) + 
  xlim(c(0,1)) +
  geom_density(size=0.75,linetype="dotted")

betaPars <- fitdistr(x=z$myVar/max(z$myVar + 0.1),start=list(shape1=1,shape2=2),"beta")
shape1ML <- betaPars$estimate["shape1"]
shape2ML <- betaPars$estimate["shape2"]

statSpecial <- stat_function(aes(x = xval, y = ..y..), fun = dbeta, colour="orchid", n = length(z$myVar), args = list(shape1=shape1ML,shape2=shape2ML))
pSpecial + statSpecial
 
```

## Questions 2 and 3
### Distribution of Brix Sugar Content of Blueberries
```{r}

# my data is read here. "$" is used to extract one column.
Berries <- read.table("BerryData.csv",header=TRUE,sep=",", stringsAsFactors=FALSE)
z <- data.frame(Berries$brix)
l <- Berries$brix
str(z)
summary(z)


# plot histogram of data
# y axis is rescaled so that the area under the curve equals one. 
p1 <- ggplot(data=z, aes(x=l, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) +  stat_bin(binwidth=1) #could add this but don't have to.
print(p1)

# add empirical density curve
# this adds a curve that id fitted to the data to smooth the shape of the histogram. It does not assume any distribution. 
p1 <-  p1 +  geom_density(linetype="dotted",size=0.75)
print(p1)

# get maximum likelihood parameters for normal
normPars <- fitdistr(Berries$brix,"normal")
print(normPars) # rpints mean and standard deviation
str(normPars)
normPars$estimate["mean"] # note structure of getting a named attribute

# plot normal probability density
# first, assign mean and sd to variables
meanML <- normPars$estimate["mean"]
sdML <- normPars$estimate["sd"]

# x values are specified from 0 to the length of variable of interest.
xval <- seq(0,max(l,na.rm=T),len=length(l))

# dnorm() is called inside ggplot's "stat_function" to generate the probability density for the normal distribution. 
stat <- stat_function(aes(x = xval, y = ..y..), fun = dnorm, colour="red", n = length(l), args = list(mean = meanML, sd = sdML))
 p1 + stat
# The red curve shows the best-fitting normal dist. Looks pretty good.
 
 
# plot exponential probability density
# fitdistr() is used to fit exponential curve to data 
expoPars <- fitdistr(l,"exponential")
rateML <- expoPars$estimate["rate"]

stat2 <- stat_function(aes(x = xval, y = ..y..), fun = dexp, colour="blue", n = length(l), args = list(rate=rateML))
 p1 + stat + stat2

# plot uniform probability density
# fitdistr() is not needed. Only two parameters are min and max of data. 
 stat3 <- stat_function(aes(x = xval, y = ..y..), fun = dunif, colour="darkgreen", n = length(l), args = list(min=min(l), max=max(l)))
 p1 + stat + stat2 + stat3

# plot gamma probability density
# fitdistr() is used again here. 
 gammaPars <- fitdistr(l,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]

stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="brown", n = length(l), args = list(shape=shapeML, rate=rateML))
 p1 + stat + stat2 + stat3 + stat4
# looks similar but better than normal dist.
 
# plot beta probability density
#it is plotted in its own plot because data must be rescaled to be between 1 and 0.
 pSpecial <- ggplot(data=z, aes(x=l/(max(l + 0.1)), y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) + 
  xlim(c(0,1)) +
  geom_density(size=0.75,linetype="dotted")

betaPars <- fitdistr(x=l/max(l + 0.1),start=list(shape1=1,shape2=2),"beta")
shape1ML <- betaPars$estimate["shape1"]
shape2ML <- betaPars$estimate["shape2"]

statSpecial <- stat_function(aes(x = xval, y = ..y..), fun = dbeta, colour="orchid", n = length(l), args = list(shape1=shape1ML,shape2=shape2ML))
pSpecial + statSpecial
# fits very well, but not a closely as gamma. 

```

## Question 4
This histogram of my data shows 4 probability density curves:
normal - red
uniform - green
exponential - blue
gamma - brown
```{r}
 p1 + stat + stat2 + stat3 + stat4
```

## Question 5
The best fit distribution for the brix content of blueberries appears to be the gamma distribution. 
the maximum likelihood parameters of this distribution are shape and rate.

The histogram profiles of the generated distribution and the original data are very similar. They both span the same range of values, and are slightly right-skewed. The original data distribution has a higher peak (about0.2) than the generated dataset (0.15), however. 

```{r}
# plot gamma probability density
# fitdistr() is used again here. 
gammaPars <- fitdistr(l,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]

stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="brown", n = length(l), args = list(shape=shapeML, rate=rateML))
p1 + stat4

# normpars() is used to find maximum likelihood parameters
normPars <- fitdistr(Berries$brix,"gamma")
print(normPars) # rpints shape and rate parameters
str(normPars)

# Make a new data set to simulate my berry data:
# rgamma() makes a random gamma distribution
# First argument is n (there were 1995 berries in my data)
# second argumant is shape
# third argument is rate
h <- rgamma(1995,20.37658929,rate=1.73798888)
g <- data.frame(1:1995,h)




# plot histogram of new data set
pnew_data <- ggplot(data=g, aes(x=h, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) +  stat_bin(binwidth=1) 
print(pnew_data)

# generated data set histogram with empirical density curve 
pnew_data <-  pnew_data +  geom_density(linetype="dotted",size=0.75)
print(pnew_data)
 

# histogram with empirical density curve of original data for comparison
p1 <-  p1 +  geom_density(linetype="dotted",size=0.75)
print(p1)
 
```
